\thetitle

### 2021-02-28 writing/results.tex
results: Add new figures with descriptions

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} using a machine with two Intel Xeon Gold 6258R processors and 500GB of memory.
Each debugging scenario had a timeout limit of 4 minutes and a
memory limit of 6GB. The complete experiment required
over 30,000 hours of compute hours.

Figure~\ref{fig:avo-matrix} summarizes our results and provides answers to
the four questions from section~\ref{sub:experiment}.  Specifically, it
depicts a head-to-head comparison of the usefulness of each mode against
every other mode.  The cell in row $R$ and column $C$ plots the percentage
of interesting debugging scenarios for which $R$ is more useful $C$ ---
we calculate this percentage as an estimated proportion from our sample.
In effect, reading across a given row $R$ provides an
overview of how $R$ compares to every other mode; the higher the bars, the
more interesting debugging scenarios for which $R$ is strictly more useful
than the other modes.  Reading down a given row $C$ provides an overview
of how every other mode compares to $C$; the higher the bars, the more
interesting debugging scenarios for which $C$ is strictly less useful than
the other modes. For example, the second cell in the row of Natural blame
shows that Natural is more useful than Natural exceptions for 12\% of all
the interesting debugging scenarios we generate from our benchmarks and
mutators.

Figure~\ref{sub:experiment} shows that the answers for questions $Q_1$
to $Q_3$ are positive. In all three semantics, blame modes outperform
their corresponding exception mode (by 9.44\% to 12.7\%). The Natural
exceptions mode is never more useful than Natural blame, and  Transient
exceptions are more useful than Transient first and Transient last blame
in 2.19\% and 2.8\% of the scenarios respectively.

For the answers to $Q_*$, we compare the percentages in antisymmetric
positions in the matrix. Blame for all three semantics is significantly
more useful than Erasure exceptions (by 33\% to 35\%). Natural blame is
more useful than both versions of Transient blame by a small percentage;
in 4.05\% of scenarios Natural blame is more useful while in 1.67\% of the
scenarios Transient blame is more useful. The Transient first and
Transient last blame are practically indistinguishable. Finally, there is
no clear winner between Natural exceptions and Transient exceptions
despite the theoretically advantageous additional checks of the Natural
semantics.

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{./plots/avo-matrix}
  \caption{Usefulness comparisons: Each cell depicts the estimated percentage of
  interesting debugging scenarios for which the row mode is more useful
  than the column mode.
  The upper bound of error margin is 0.02\%}
  \label{fig:avo-matrix}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{./plots/bt-lengths-table}
  \caption{Programmer effort: Each plot depicts the distribution of trail
  lengths for a given mode across all benchmarks.}
  \label{fig:effort-table}
\end{figure}

Figure~\ref{fig:effort-table} shows the distribution of programmer effort
for our sample of interesting debugging scenarios. Unlike the usefulness
comparisons above, these proportions do not generalize to a representation of the full population.
Still, as we discuss in section~\ref{subsec:effort}, they provide an
alternative view of the workings of the rational programmer.

There are
three immediate take-aways from the figure. First, the effort for successfully
debugging interesting scenarios (in green) for the random mode of the
rational programmer follows  a normal distribution, as expected.\footnote{The
random mode distribution is indistinguishable for all three semantics and thus the figure~\ref{fig:effort-table} 
only shows one plot.} In contrast, in the other modes, successful effort coalesces at
the left side of the plot, meaning that in most cases the programmer needs
to type a single component to debug a scenario.

A third notable piece of information from
the figure is that every mode has failed debugging scenarios, not just
Erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program, or only typed modules. This situation
can even happen immediately, which explains failures with 0 effort.

@@

>>>>

We run the experimental process from section~\ref{sec:rational} on the
sampled debugging scenarios from section~\ref{sec:sample} using a machine with two Intel Xeon Gold 6258R processors (28 doubly-threaded cores each) and 500GB of memory.
Each debugging scenario had a timeout limit of 4 minutes and a
memory limit of 6GB. The complete experiment required
over 30,000 hours of compute hours.

Figure~\ref{fig:success-bars} summarizes the overall success rates of every mode.
The success rates illustrate a few points that underly the rest of our analysis.
The first notable piece of information from
this figure is that every mode has failed debugging scenarios, not just
Erasure. This should
not come as a surprise to the astute
reader. After all, as we
discuss above, some modes are more successful than others in certain
scenarios. Indeed any mode can fail if running a scenario results in an
exception and the exception carries no useful information about which
module the programmer should type next, e.g., because the stack trace 
does not contain frames from any module of the program, or only typed modules. This situation
can happen several steps along a blame trail or even immediately (figure~\ref{fig:effort-table} dives into this below).

@@

The second key observation from figure~\ref{fig:success-bars} is that the modes that use blame all outperform those that do not.
In particular, Natural and both of Transient's blame modes succeed in around 95\% of our scenarios, while their corresponding exception modes succeed in around 85\% of them and Erasure succeeds in only 60\% of them.
The only exception is that the Random programmer always succeeds;
this reflects the fact that every scenario has finitely many modules and this mode can never get stuck.

Figure~\ref{fig:avo-bars} depicts a head-to-head comparison of every mode's performance against every other mode, helping to answer the four questions from section~\ref{sub:experiment}.
Each plot shows a breakdown of the proportion of scenarios where one mode performs better or worse than each other mode.
In particular, each bar above zero represents the proportion where the plot's entitled mode succeeds and the mode on the x-axis fails; the correspnding bar below zero represents the proportion of the opposite case.
For example, the plot titled ``Natural'' shows that Natural outperforms Natural exceptions in about 10\% of the scenarios, and the inverse (Natural performs wrse than Natural exceptions) never happens.
Similarly, the plot titled ``Transient last blame'' shows that Transient last blame outperforms Natural exceptions about in abut 10\% of the scenarios, but conversely it performs worse than Natural exceptions in about 2\% of the scenarios.

Figure~\ref{fig:avo-bars} shows that the answers for questions $Q_1$ to $Q_3$ are all positive.
In all three semantics, blame modes outperform their corresponding exception mode (by around 10\% for each one).
The Natural exceptions mode is never more useful than Natural blame, and  Transient
exceptions are more useful than Transient first and Transient last blame
in a small percentage (about 2\%) of the scenarios.

Figure~\ref{fig:avo-bars} also provides answers to $Q_*$.
Blame for all three semantics is significantly
more useful than Erasure exceptions (by around 30\%). Natural blame is
more useful than both versions of Transient blame by a small percentage;
in about 4\% of scenarios Natural blame is more useful while in about 1.5\% of the
scenarios Transient blame is more useful. The Transient first and
Transient last blame are practically indistinguishable. Finally, there is
no clear winner between Natural exceptions and Transient exceptions
despite the theoretically advantageous additional checks of the Natural
semantics.

In order to better understand questions $Q_1$ to $Q_3$, figure~\ref{fig:success-venns} depicts a breakdown of the success of each semantics in comparison to Erasure.
Specifically, the figure shows one Venn diagram per mode of the rational programmer that uses blame.
Each diagram breaks breaks down the overlap in the success of the blame mode, its corresponding exception mode, and Erasure.
For example, in the breakdown of Natural, the diagram shows that all three modes succeed on the same scenario 57.3\% of the time, only Natural and Natural exceptions succeed on 29.1\% of the scenarios, only Natural and Erasure succeed on 2.1\%, and Natural alone succeeds on 9\%.
This analysis highlights the success tradeoffs each semantics offers against Erasure, with and without blame, to answer questions about the tradeoffs in choosing to use blame or just exceptions for each semantics.
For instance, the breakdown of Natural clearly illustrates that if choosing between Natural blame, Natural exceptions, and Erasure, Natural blame is the absolutely most successful: all of the successes of the other two modes are subsets of Natural's successful scenarios.
On the other hand, Transient's blame modes are similar but not so clear-cut.

\begin{figure}
  \centering
  \includegraphics[width=0.30\textwidth]{./plots/success-bars}
  \caption{Success baselines: The estimated percentage of scenarios for which each mode succeeds in locating the bug.
  }
  \label{fig:success-bars}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{./plots/avo-bars}
  \caption{Usefulness comparisons: Each plot depicts a head-to-head comparison of the entitled mode against every other mode.
  The portion above 0 (in green) is the estimated percentage of scenarios where the entitled mode is more useful than each of the other modes.
  The portion below 0 (in red) is the estimated percentage of scenarios with the inverse situation.
  The upper bound of error margin is 0.02\%.}
  \label{fig:avo-bars}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.32\textwidth]{./plots/TR-TR-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-newest-transient-stack-first-venn}
  \hfill
  \includegraphics[width=0.32\textwidth]{./plots/transient-oldest-transient-stack-first-venn}

  \caption{Blame usefulness analysis: Each diagram shows the breakdown of how three modes' successful scenarios overlap.
  The percentage of scenarios covered by each circle are proportions for which the corresponding mode succeeds.
  For example, in the breakdown of Natural, all three modes succeed on the same scenario 57.3\% of the time, only Natural and Natural exceptions succeed on 29.1\% of the scenarios, only Natural and Erasure succeed on 2.1\%, and Natural alone succeeds on 9\%.
  }
  \label{fig:success-venns}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./plots/bt-lengths-table}
  \caption{Programmer effort: Each plot depicts the distribution of trail
  lengths for a given mode across all benchmarks.}
  \label{fig:effort-table}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{./plots/bt-length-comparisons}
  \caption{Effort comparisons: Each plot depicts the distribution of scenarios with trail length differences ranging from -3 to 3.
  A negative difference means that the top mode's trail is three steps shorter than the bottom mode's trail for the same scenario, and a positive difference is the reverse.
  Only scenarios where both modes succeed are considered.
  The \texttt{.2|.8} on the y-axis indicates that the axis is truncated between 0.2 and 0.8.}
  \label{fig:effort-comparisons}
\end{figure}

Turning to programmer effort, figure~\ref{fig:effort-table} shows the distribution of blame trail lengths
for our sample of interesting debugging scenarios. Unlike the usefulness
comparisons above, these proportions do not generalize to a representation of the full population.
Still, as we discuss in section~\ref{subsec:effort}, they provide an
alternative view of the workings of the rational programmer.

There are
two immediate take-aways from the figure. First, the effort for successfully
debugging interesting scenarios (in green) for the random mode of the
rational programmer follows  a normal distribution, as expected.\footnote{The
random mode distribution is indistinguishable for all three semantics and thus the figure~\ref{fig:effort-table} 
only shows one plot.} In contrast, in the other modes, successful effort coalesces at
the left side of the plot, meaning that in most cases the programmer needs
to type a single component to debug a scenario.

Figure~\ref{fig:effort-comparisons} provides head-to-head comparisons of effort.
To directly compare effort between modes, we compute the difference in trail length between two modes for all scenarios where they both succeed.
Hence, each plot in the figure shows the distribution of scenarios with length differences ranging from -3 (the first mode's trail is 3 steps shorter than the second) to 3 (the first mode's trail is 3 steps longer than the second).
The figure offers several insights about the comparative effort for each mode that complement the insights about comparitive success from figure~\ref{fig:avo-bars}.
First, Natural blame never produces shorter trails than Natural exceptions, and sometimes (albeit rarely) produces slightly longer ones.
Second, Natural relatively often (around 10\% of the scenarios) produces shorter trails than both Transient blame modes, and they are sometimes significantly shorter.
Third, Transient last blame often produces shorter trails than Transient first blame, which is a particularly interesting distinction in light of their nearly undistinguishable success rates.
Finally, Transient's blame modes also share the characteristic with Natural that blame sometimes produces longer trails than their corresponding exception modes.
